---
title: "R Profiler"
output:
  html_document:
    df_print: paged
---



```{r}
plot(cars)
```

# Tools for R Profiler
##Profiling

## Is your code running slowly?
    
### First you should think about, is how to get your code to run and make it readble

#### - Design first, then optimize
#### - Remember: Premature optimization is the root of all evil.
#### - Measure (collect data), don guess.
#### - If you're going to be a scientist you need to apply the same principles here!

system.time()

##### Computes the time in seconds needed to execute an expression

######      - If there's an error, gives time until the error occurred

##### Returns an object of class proc_time

###### - user time: time charged to the CPU(s) for this expression
###### - elapsed time: "Wall clock" time

Elapsed time may be smaller than user time if the machine has multi processors/cores
Elapsed time may be greater than user time if the CPU spends a lot of time waiting arround


Multi-threaded  BLAS libraries take advantage of multi core processors

Parallel processing via the parallel package, use multi computers, or something like computer clusters.

Example:

```{r}
system.time(readLines("http://www.jhsph.edu"))

hilbert <- function(n){
    
    i <- 1:n
    1/outer(i-1, i, "+")
}

x <- hilbert(1000)
system.time(svd(x))


```


The Rprof function starts the profiler in R
 - R must be compiled with profiler support
The summaryRprof() function summarizes the output from Rprof() (otherwise itś not readble)
DO NOT use system.time() and Rprof() fucntions together... or sad you will be... kkk








