---
title: "Generating Random Numbers"
author: "José Sousa"
output:
  pdf_document: default
  html_document:
    df_print: paged
    smart: no
header-includes: \usepackage{bbm}
---
### Some of the functions in R that can be use to do Simulation:

* Functions to generate random variables
    + rnorm()
    + dnorm()
    + pnorm()
    + rpois()

#### Probability distribution functions usually have 4 distribution functions associated with them. The functions prefixed with a 

    d for density
    r for random number generation
    p for cumulative distribution
    q for quantile function


#### The set.seed() function allows us to reproduce the sequence of random variables that was generated for the given seed
```{r}
set.seed(1)
rnorm(5)

set.seed(2)
rnorm(5)
```

##### A new random simulation without the seed, has you will see, it will be difficult to reproduce this random sequence.

```{r}
rnorm(5)
```

#### Now for the first two simulations, we can reproduce the exact seequence of random variables:

##### Reproducing second sequence, seed 2

```{r}
set.seed(2)
rnorm(5)
```

##### Reproducing second sequence, seed 1

```{r}
set.seed(1)
rnorm(5)
```

# Simulating Random numbers for a linear model

### Suppose we want to simulate from the follwing linear model 

#### $y = \beta{_0}+\beta{_1}\mathrm{x}+\varepsilon$

#### where $\varepsilon \sim N(0,1{_2})$. Assume $\mathrm{x}\sim \mathrm{N}(0,1{_2}), \beta{_0} = 0.5$ and $\beta{_1} = 2$






```{r}
set.seed(20)
x <- rnorm(100, 1, 2)
e <- rnorm(100, 0, 1)
y <- 0.5 + 2 * x + e
summary(y)
plot(x,y)
```

### x as binomial variable
    What if x is binary
    
```{r}
set.seed(10)
x <- rbinom(100, 1, 0.5)
e <- rnorm(100, 0, 2)
y <- 0.5 + 2 * x + e
summary(y)
plot(x,y)
```


## Generating Random Numbers From a Generalized Linear Model

#### Suppose we want to simulate from a Poisson model where

$Y \sim Poisson(\mu)$

$\log \mu = \beta{_0} + \beta{_1}x$

#### and $\beta{_0} = 0.5$ and $\beta{_1} = 0.3$. We need to use rpois function for this

```{r}
set.seed(1)
x <- rnorm(100)
log.mu <- 0.5 + 0.3 * x
y <- rpois(100, exp(log.mu))
summary(y)
plot(x, y)

```
```{r}


```